Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-08 paper.tex
Updates & edits.

% tell a nell story up front
Miles from the nearest school, a young girl in Ethiopia turns on a
tablet computer for the first time.  The solar-powered machine begins
talking to her immediately: \textit{``Hello!  Would you like to hear a story?''}

She listens intently to a story about a princess named Nell.  Later, when
the girl has learned a little more, she will give the princess her own
name, ``Rahel,'' and change all the colors, but for now the green
book draws pictures for her and asks her to trace shapes on the screen.

@@

As she traces the R, it comes to life and gallops across the screen.
\textit{``Run starts with R.  Roger the R Runs across the Red Rug.  He has a
dog named Rover.''}  Rover barks: \textit{``Ruff! Ruff!''}  The Princess asks,
\textit{``Can you find something Red?''} and Rahel uses the camera to
photograph her red boots.

@@

As Rahel grows, the book asks her to trace not just letters, but whole
words.  The book's responses are written on the screen as it speaks
them, and eventually she doesn't need to leave the sound on all the
time.  Soon Rahel can write complete sentences in her special book,
and the Princess will respond sometimes.  New stories teach her about
music (by playing certain tunes she unlocks a dungeon door) and
programming with blocks (a not-very-bright turtle needs Princess Rahel's
help to draw different shapes).  Rahel writes her own stories,
which she shares with her friends.

An older Rahel learns that the block language she used to talk with
the turtle is also used to write all the software running inside
her special book.  The stories Rahel writes can now include new
functionality that make her book even more exciting for the kids
just now opening their green meta books.

@@

We believe the combination of these four key ideas provides a novel
experience which addresses some of the shortcomings of earlier
learning environments for children.  In this section we will discuss
our motivating concepts, and in the next we will describe how
they are implemented in Nell.

@@

Children (like all humans) are hard-wired for
stories~\cite{boyd:stories}.  Nell uses a
\textit{Narrative Interface}~\cite{bizzocchi:narrative,don:narrative,laurel:computers} 
to guide the child user through the pedagogical capabilities of the system.

% excerpt of story model
% not always a story -- kid can draw on their own and return to the
% story; Nell will observe the free play and offer suggestions.
\subsection{Personal}
% customization, renaming characters
\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% it has a ``low floor and no ceiling'' to allow a child to master more
% and more of the system as she grows.

>>>>

% tell a nell story up front
Miles from the nearest school, a young Ethiopian girl turns on a
tablet computer for the first time.  The solar-powered machine begins
talking to her immediately: \textit{``Hello!  Would you like to hear a story?''}

She nods and listens to a story about a princess.  Later, when
the girl has learned a little more, she will tell the machine that the
princess is named ``Rahel,'' like she is, and change all its
colors---but for now the green book draws pictures of the unnamed Princess
for her and asks her to trace shapes on the screen.

@@

As she traces the R, it comes to life and gallops across the screen.
\textit{``Run starts with R.  Roger the R runs across the Red Rug.  Roger has a
dog named Rover.''}  Rover barks: \textit{``Ruff! Ruff!''}  The Princess asks,
\textit{``Can you find something Red?''} and Rahel uses the camera to
photograph her new boots.

@@

As Rahel grows, the book asks her to trace not just letters, but whole
words.  The book's responses are written on the screen as it speaks
them, and eventually she doesn't need to leave the sound on all the
time.  Soon Rahel can write complete sentences in her special book,
and sometimes the Princess will respond to them.  New stories teach her about
music (she unlocks a dungeon door by playing certain tunes) and
programming with blocks (Princess Rahel helps a not-very-bright turtle
to draw different shapes).  Rahel writes her own stories about the Princess,
which she shares with her friends.

An older Rahel learns that the block language she used to talk with
the turtle is also used to write all the software running inside
her special book.  The stories Rahel writes can now include new
functionality that make her book even more exciting for the kids
just now receiving their own green meta books.

@@

We believe the combination of these four key concepts provides a novel
experience which addresses some of the shortcomings of earlier
learning environments for children.  In this section we will discuss
these four building blocks, and in the next section we will describe their
implementation in Nell.

@@

Children (like all humans) are hard-wired for
stories~\cite{boyd:stories}.  Nell uses a
\textit{Narrative Interface}~\cite{bizzocchi:narrative,don:narrative,laurel:computers} 
to guide the child user through the pedagogical capabilities of the
system.  All system actions are shaped by the storybook metaphor, and
interactions with the system are conceptualized as interactions with
one of the system protagonists.

Nell's overall story is a multi-character serial adventure, taking
cues from \textit{The Diamond Age} and
from the UNIVERSE narrative-generation system \cite{lebowitz:universe84}.
Each of the several characters represents a specific skill or subject
area, and each adventure represents about a year's curriculum.  This
design also improves modularity and allows updates and decentralized
authorship.

As illustrated in figure~\ref{fig:nell}, Nell's characters are
always-available agents running on top of a particular system \textit{activity}
(application) which provides specialized functionality.  The agents
are not always foregrounded: constructionist learning occurs when the
child plays freely to ``make things'' with the base activity.
The handwriting tutor is fundamentally a drawing activity; the
adventure involving the magical musical lock is also a music-making activity.
%  ``Making things'' is the key to constructionist learning.
The narrative system is
hooked into each activity to provide passive guidance
(congratulating the child when it notices they've drawn a letter),
active guidance (Apple-Guide--style~\cite{powers:appleguide}
contextual help), or system services (switch activities; jump into a
related story module/lesson plan).

% Story Module == Lesson Plan

% not always a story -- kid can draw on their own and return to the
% story; Nell will observe the free play and offer suggestions.
\subsection{Personalization}
% customization, renaming characters

Children often decorate their belongings, and the
child owning a Nell system is encouraged to personalize it.
Superficially, colors can be changed and characters can be renamed and
dressed up.  But Nell also supports deeper story personalization.

Choices the child has made are stored in \textit{Avatar Services},
accessible to all story modules and activities.  Past choices
can thus influence the future direction of a story.  Story module
alternatives are supported to allow learning material to presented in
different ways for different learners. Further, the record of a
child's experiences with Nell are recorded in a \textit{Journal},
which can be edited and customized by the user.  Careful arrangement
of sections of the Journal can become a Portfolio~\cite{stefanakis:portfolios}.
The journal also allows undo/redo \textbf{XXXX}.

In the future, we expect to allow further personalization via deep
recombination of story elements.  Lebowitz~\cite{lebowitz:universe85} and Riedl~\cite{riedl:planning} show how a planner can be used to recombine story
modules.  We expect to be less ambitious then previous work: instead
of attempting to generate thousands of stories from tens of templates,
we expect to modestly adapt from hundreds of story modules created
in a decentralized manner by teachers.

\subsection{Direct interaction}
% constructionist?
% handwriting as UI?
% motor skills
% photo of XO-3?

\subsection{Growable (``no ceiling'')}
% it has a ``low floor and no ceiling'' to allow a child to master more
% and more of the system as she grows.
